---
title: "Appalachia_SHPs"
author: "Jared Gingrich"
date: "4/6/2022"
output: html_document
---

This script subsets US Census Bureau TIGER/Line shapefiles for counties and smaller geographic units that make up Appalachia

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      warning = F,
                      message = F)

#Loading packages from the package script
source("./Scripts/load_pkg.R", local = knitr::knit_global())
```

Appalachian counties, data from the [Appalachia Regional Commission](https://www.arc.gov/map/subregions-in-appalachia/)

```{r}
appalachia <- read_excel("./Data/Subregions_2009_Data-1.xls", sheet = 1, skip=3)
```

US Census Bureau TIGER/Line Census Tract shapefiles, subset for Appalachia. Data [here](https://www2.census.gov/geo/tiger/TIGER2019/TRACT/)

* Combining census tracts shapefiles for the 13 states that are part of Appalachia performed in QGIS
* Adjusting FIPS to reflect combined Virginia City-County units

```{r}
app_tract <- read_sf("./Data/Census/tl_2019_tract.shp", quiet=T) #Loading Appalachia census tract shapefile from QGIS
app_tract <- app_tract %>%
  mutate(FIPS = paste0(STATEFP, COUNTYFP)) %>% #Creating new FIPS code field
  mutate(FIPS = ifelse(FIPS == "51520", "51191", FIPS)) %>% #Bristol City
  mutate(FIPS = ifelse(FIPS == "51530", "51163", FIPS)) %>% #Buena Vista City
  mutate(FIPS = ifelse(FIPS == "51580", "51005", FIPS)) %>% #Covington City
  mutate(FIPS = ifelse(FIPS == "51640", "51035", FIPS)) %>% #Galax City
  mutate(FIPS = ifelse(FIPS == "51678", "51163", FIPS)) %>% #Lexington City
  mutate(FIPS = ifelse(FIPS == "51690", "51089", FIPS)) %>% #Martinsville City
  mutate(FIPS = ifelse(FIPS == "51720", "51195", FIPS)) %>% #Norton City
  mutate(FIPS = ifelse(FIPS == "51750", "51121", FIPS)) %>% #Radford City 
  filter(FIPS %in% appalachia$FIPS) #filtering for just counties in Appalachia
app_tract <- st_transform(app_tract, st_crs(5070)) #projection 5070: Conus Albers
```

US Census Bureau TIGER/Line Urban Area shapefiles, subset for Appalachia. Data [here](https://www2.census.gov/geo/tiger/TIGER2010/UA/2010/)

* Splitting Appalachian urban areas by census block performed in QGIS
* Calculating rural and urban population counts at the census tract level

```{r}
ua_blocks <- read_sf("./Data/Census/Urban Area Blocks.shp", quiet=T)

ua_tract <- st_drop_geometry(ua_blocks) %>%
  mutate(TRACTID = substr(BLOCKID10, 1, 11)) %>% #Creating census tract FIPS codes
  group_by(TRACTID) %>% #grouping by census tract
  summarize(Total_POP = sum(POP10), #summarizing rural and urban population counts
            Urban_POP = sum(POP10[Urban==1]),
            Rural_POP = sum(POP10[Urban==0])) %>%
  mutate(PctRURAL = Rural_POP/Total_POP*100) %>% #percent rural and urban
  mutate(RURAL = ifelse(PctRURAL > 50, "Rural", "Urban")) #majority population
write.csv(ua_tract, "./Out/Rural_Urban_Tracts.csv")

#Merging with Appalachian census tracts
app_tract <- merge(app_tract, ua_tract, by.x="GEOID", by.y="TRACTID")

write_sf(app_tract, "./Out/SHPs/Appalachia_Tracts.shp")
```

US Census Bureau TIGER/Line State shapefile. Data [here](https://www2.census.gov/geo/tiger/GENZ2019/shp/)

* Subsetting states in the continental US 
* Subsetting states that include portions of Appalachia

```{r warning=F, message=F}
state <- read_sf("./Data/Census/cb_2019_us_state_20m.shp", quiet=T) %>%
  filter(!NAME %in% c("Puerto Rico", "Hawaii", "Alaska")) #continental US
state <- st_transform(state, st_crs(5070)) #projection 5070: Conus Albers
write_sf(state, "./Out/SHPs/CONUS_States.shp", quiet=T)

app_state <- state %>% #Appalachia states
  filter(STUSPS %in% c("AL", "GA", "KY", "MD", "MS", "NC", "NY", "OH", "PA", "SC", "TN", "VA", "WV"))
write_sf(app_state, "./Out/SHPs/Appalachia_States.shp", quiet=T)
```
