---
title: "Indicators of Access"
author: "Jared Gingrich"
date: "4/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      warning = F,
                      message = F)

#Loading packages from the package script
source("./Scripts/load_pkg.R", local = knitr::knit_global())

#Loading script for the ACS 5 year estimate data function
source("./Scripts/acs5y.R", local = knitr::knit_global())
```

# Distance-to-Store

This script joins store locations from the SNAP retailer database to the historic SNAP retailer database.

* SNAP retailer database has the most up-to-date information, historic SNAP database has store classifications.
* SNAP retailer database downloaded [here](https://usda-fns.hub.arcgis.com/datasets/USDA-FNS::snap-store-locations?geometry=124.057%2C-9.799%2C-153.326%2C74.211).
* Historic SNAP retailer database downloaded [here](https://www.fns.usda.gov/snap/retailer/historicaldata).

```{r}
#Joining stores with Retailer Policy and Management table store categorizations

stores <- read_sf("./Data/Distance-to-Store/Appalachia_Store_Locations.shp", quiet=T)
hist <- st_drop_geometry(read_sf("./Data/Distance-to-Store/Historical SNAP Retailer Locator.shp", quiet=T))

st_hist <- merge(st_drop_geometry(stores) %>%
                   mutate(Latitude = round(Latitude, digits=4),
                          Longitude = round(Longitude, digits=4)), hist %>%
                   mutate(Latitude = round(as.numeric(Latitude), digits=4),
                          Longitude = round(as.numeric(Longitude), digits=4),
                          End_Year = sapply(strsplit(End.Date, "/"), '[', 3)) %>%
                   dplyr::select(Store.Name, Latitude, Longitude, Store.Type, End_Year), 
                 by.x=c("Latitude", "Longitude"),
                 by.y=c("Latitude", "Longitude"))
st_hist <- st_hist %>%
  mutate(End_Year = ifelse(is.na(End_Year), 2021, End_Year)) %>%
  group_by(ObjectId) %>%
  mutate(Max_End_Year = max(End_Year)) %>%
  ungroup() %>%
  filter(End_Year == Max_End_Year) %>%
  group_by(ObjectId) %>%
  summarize(Count=n(),
            Store_Name = unique(Store_Name),
            TYPE = unique(TYPE),
            Store_Type = paste(unique(Store.Type), collapse=";"),
            Store_Type2 = paste(sort(unique(Store.Type[End_Year == Max_End_Year])), collapse=";"))

st_groc <- st_hist %>%
  filter(TYPE == "Grocery") %>%
  mutate(Type1 = sapply(strsplit(Store_Type2, ";"), '[', 1),
         Type2 = sapply(strsplit(Store_Type2, ";"), '[', 2),
         Type3 = sapply(strsplit(Store_Type2, ";"), '[', 3)) %>%
  mutate(Store_Type = ifelse(Type1 %in% c("Supermarket", "Super Store", "Large Grocery Store"), "Grocery", NA),
         Store_Type = ifelse(Type2 %in% c("Supermarket", "Super Store", "Large Grocery Store"), "Grocery", Store_Type),
         Store_Type = ifelse(Type3 %in% c("Supermarket", "Super Store", "Large Grocery Store"), "Grocery", Store_Type),
         Store_Type = ifelse(is.na(Store_Type), "Other", "Grocery"))

stores <- merge(stores %>%
                  filter(TYPE == "Grocery"), 
                st_groc %>%
                  dplyr::select(ObjectId, Store_Type), by="ObjectId", all=T)
missing <- stores %>%
  filter(is.na(Store_Type))
missing <- merge(missing %>%
                   mutate(Latitude = round(Latitude, digits=5),
                          Longitude = round(Longitude, digits=5)), 
                 hist %>%
                   mutate(Latitude = as.numeric(Latitude),
                          Longitude = as.numeric(Longitude),
                          End_Year = sapply(strsplit(End.Date, "/"), '[', 3)) %>%
                   dplyr::select(Store.Name, Latitude, Longitude, Store.Type, End_Year),
                 by=c("Latitude", "Longitude"))
missing <- st_drop_geometry(missing) %>%
  mutate(End_Year = ifelse(is.na(End_Year), 2021, End_Year)) %>%
  group_by(ObjectId) %>%
  mutate(Max_End_Year = max(End_Year)) %>%
  ungroup() %>%
  filter(End_Year == Max_End_Year) %>%
  group_by(ObjectId) %>%
  summarize(Count=n(),
            Store_Name = unique(Store_Name),
            TYPE = unique(TYPE),
            Store_Type = paste(unique(Store.Type), collapse=";"),
            Store_Type2 = paste(sort(unique(Store.Type[End_Year == Max_End_Year])), collapse=";"))
missing <- missing %>%
  filter(TYPE == "Grocery") %>%
  mutate(Type1 = sapply(strsplit(Store_Type2, ";"), '[', 1),
         Type2 = sapply(strsplit(Store_Type2, ";"), '[', 2),
         Type3 = sapply(strsplit(Store_Type2, ";"), '[', 3)) %>%
  mutate(Store_Type = ifelse(Type1 %in% c("Supermarket", "Super Store", "Large Grocery Store"), "Grocery", NA),
         Store_Type = ifelse(Type2 %in% c("Supermarket", "Super Store", "Large Grocery Store"), "Grocery", Store_Type),
         Store_Type = ifelse(Type3 %in% c("Supermarket", "Super Store", "Large Grocery Store"), "Grocery", Store_Type),
         Store_Type = ifelse(is.na(Store_Type), "Other", "Grocery"))
stores <- merge(stores,
                missing %>%
                  mutate(Store_Type_missing = Store_Type) %>%
                  dplyr::select(ObjectId, Store_Type_missing), by="ObjectId", all=T)
stores <- stores %>%
  mutate(Store_Type = ifelse(is.na(Store_Type), Store_Type_missing, Store_Type))

missing <- stores %>%
  filter(is.na(Store_Type))

missing <- merge(missing %>%
                   mutate(Address2 = tolower(Address)), 
                 hist %>%
                   mutate(Address2 = tolower(paste0(Street.Num, " ", Street.Nam)),
                          End_Year = sapply(strsplit(End.Date, "/"), '[', 3)) %>%
                   dplyr::select(Store.Name, Address2, Zip.Code, Store.Type, End_Year),
                 by.x=c("Zip5", "Address2"), by.y=c("Zip.Code", "Address2"))
missing <- st_drop_geometry(missing) %>%
  mutate(End_Year = ifelse(is.na(End_Year), 2021, End_Year)) %>%
  group_by(ObjectId) %>%
  mutate(Max_End_Year = max(End_Year)) %>%
  ungroup() %>%
  filter(End_Year == Max_End_Year) %>%
  group_by(ObjectId) %>%
  summarize(Count=n(),
            Store_Name = unique(Store_Name),
            TYPE = unique(TYPE),
            Store_Type = paste(unique(Store.Type), collapse=";"),
            Store_Type2 = paste(sort(unique(Store.Type[End_Year == Max_End_Year])), collapse=";"))
missing <- missing %>%
  filter(TYPE == "Grocery") %>%
  mutate(Type1 = sapply(strsplit(Store_Type2, ";"), '[', 1),
         Type2 = sapply(strsplit(Store_Type2, ";"), '[', 2),
         Type3 = sapply(strsplit(Store_Type2, ";"), '[', 3)) %>%
  mutate(Store_Type = ifelse(Type1 %in% c("Supermarket", "Super Store", "Large Grocery Store"), "Grocery", NA),
         Store_Type = ifelse(Type2 %in% c("Supermarket", "Super Store", "Large Grocery Store"), "Grocery", Store_Type),
         Store_Type = ifelse(Type3 %in% c("Supermarket", "Super Store", "Large Grocery Store"), "Grocery", Store_Type),
         Store_Type = ifelse(is.na(Store_Type), "Other", "Grocery"))
stores <- merge(stores %>%
                  dplyr::select(-Store_Type_missing),
                missing %>%
                  mutate(Store_Type_missing = Store_Type) %>%
                  dplyr::select(ObjectId, Store_Type_missing), by="ObjectId", all=T)
stores <- stores %>%
  mutate(Store_Type = ifelse(is.na(Store_Type), Store_Type_missing, Store_Type))

stores <- stores %>%
  dplyr::select(-Store_Type_missing)

missing <- stores %>%
  filter(is.na(Store_Type))
#Remaining 31 for manual fixing
write_sf(stores, "./Data/Distance-to-Store/Appalachia_Store_Grocery.shp", quiet=T)
```

This script calculates the distance-to-store indicator of access (distance to the nearest store) at the census tract level.

* Distances to the nearest store for census blocks split along a regular 500-meter-square grid were computed in QGIS, using the Distance to Nearest Hub tool. 
* Distances are computed three times, using the locations of only large grocery stores, all grocery stores, and grocery stores and farmers markets. 
* Census block shapefiles with population and housing counts downloaded [here](https://www2.census.gov/geo/tiger/TIGER2010BLKPOPHU/).

```{r}
#aggregating from the rectangular grid to the census blocks
grid <- read.csv("./Data/Distance-to-Store/Grid_Store_Distances.csv")
grid <- grid %>%
  mutate(BLOCKID = ifelse(nchar(as.character(BLOCKID10)) == 14, paste0("0", as.character(BLOCKID10)), as.character(BLOCKID10))) %>%
  mutate(FIPS = substr(BLOCKID, 1, 5))

#calculating closest grocery or farmers market distance
grid <- grid %>%
  mutate(GrocFMDist = ifelse(Groc_Dist <= FM_Dist, Groc_Dist, FM_Dist))

#as always, fixing the Virginia city-county discrepancies
grid <- grid %>%
  mutate(FIPS = ifelse(FIPS == "51520", "51191", FIPS)) %>% #Bristol City
  mutate(FIPS = ifelse(FIPS == "51530", "51163", FIPS)) %>% #Buena Vista City
  mutate(FIPS = ifelse(FIPS == "51580", "51005", FIPS)) %>% #Covington City
  mutate(FIPS = ifelse(FIPS == "51640", "51035", FIPS)) %>% #Galax City
  mutate(FIPS = ifelse(FIPS == "51678", "51163", FIPS)) %>% #Lexington City
  mutate(FIPS = ifelse(FIPS == "51690", "51089", FIPS)) %>% #Martinsville City
  mutate(FIPS = ifelse(FIPS == "51720", "51195", FIPS)) %>% #Norton City
  mutate(FIPS = ifelse(FIPS == "51750", "51121", FIPS)) #Radford City

#dropping non-Appalachia edge block fragments
appalachia <- read_excel("./Data/Subregions_2009_Data-1.xls", sheet = 1, skip=3)
grid <- grid %>% filter(FIPS %in% appalachia$FIPS)

#FARA Low Access Populations
grid <- grid %>%
  mutate(LApop_05 = ifelse(Large_Dist > 0.5, SPLIT_POP, 0),
         LApop_1 = ifelse(Large_Dist > 1, SPLIT_POP, 0),
         LApop_10 = ifelse(Large_Dist > 10, SPLIT_POP, 0),
         LApop_20 = ifelse(Large_Dist > 20, SPLIT_POP, 0))

#Aggregating grid squares to census blocks
app_block <- grid %>%
  group_by(BLOCKID) %>%
  summarize(POP = mean(POP10),
            Large_Dist = ifelse(POP != 0, sum(SPLIT_POP*Large_Dist)/POP, 0),
            Groc_Dist = ifelse(POP != 0, sum(SPLIT_POP*Groc_Dist)/POP, 0),
            GrocFMDist = ifelse(POP != 0, sum(SPLIT_POP*GrocFMDist)/POP, 0),
            LApop_05 = round(sum(LApop_05)),
            LApop_1 = round(sum(LApop_1)),
            LApop_10 = round(sum(LApop_10)),
            LApop_20 = round(sum(LApop_20)))

#Aggregating census blocks to census tracts
#and fixing FIPS codes for Virginia County-City combinations
app_tract <- app_block %>%
  mutate(TRACTID = substr(BLOCKID, 1, 11)) %>%
  mutate(FIPS = substr(BLOCKID, 1, 5)) %>%
  mutate(FIPS = ifelse(FIPS == "51520", "51191", FIPS)) %>% #Bristol City
  mutate(FIPS = ifelse(FIPS == "51530", "51163", FIPS)) %>% #Buena Vista City
  mutate(FIPS = ifelse(FIPS == "51580", "51005", FIPS)) %>% #Covington City
  mutate(FIPS = ifelse(FIPS == "51640", "51035", FIPS)) %>% #Galax City
  mutate(FIPS = ifelse(FIPS == "51678", "51163", FIPS)) %>% #Lexington City
  mutate(FIPS = ifelse(FIPS == "51690", "51089", FIPS)) %>% #Martinsville City
  mutate(FIPS = ifelse(FIPS == "51720", "51195", FIPS)) %>% #Norton City
  mutate(FIPS = ifelse(FIPS == "51750", "51121", FIPS)) %>% #Radford City
  group_by(TRACTID) %>%
  summarize(TOT_POP = sum(POP),
            Large_Dist = sum(Large_Dist*POP, na.rm=T)/TOT_POP,
            Groc_Dist = sum(Groc_Dist*POP, na.rm=T)/TOT_POP,
            GrocFMDist = sum(GrocFMDist*POP, na.rm=T)/TOT_POP,
            LApop_05 = sum(LApop_05), 
            LApop_1 = sum(LApop_1),
            LApop_10 = sum(LApop_10),
            LApop_20 = sum(LApop_20))

write.csv(app_tract, "./Out/Indicators_of_access/Appalachia_Tract_Distances.csv")
```

# Economic- Median Household Income

This script calculates Median Household Income at the census tract level, as a proportion of US Median Household Income and adjusted for the Bureau of Economic Regional Purchasing Parities. Median incomes for different household sizes are calculated and averaged to account for household structure, similar to the calculation of poverty rates for different household categories. Some census tract median household incomes are redacted, and these missing values are replaced by the corresponding county-level Median Household Income proportions. Two tables are returned, one for Appalachian census tracts alone, one for all US census tracts.

American Community Survey, 2015-2019 estimates

* ACS Table B19019: Median household income by household size
* ACS Table B11016: Household type and size
* ACS Table B01003: Population

Bureau of Economic Analysis

* Metropolitan Area Regional Price Parities (MARPP)
* Portion Area Regional Price Parities (MARPP)

```{r}
#Determining county-level regional price parities
#Metropolitan statistical area counties
msa <- read_excel("./Data/Economic/list1_2020.xls", skip = 2)
colnames(msa) <- c("CBSA_FIPS", "Metro_Div", "CSA_Code", "CBSA_Title", "Met_Mic", "Div_Title", "CSA_Title", "COUNTY", "STATE", "ST_FIPS", "CTY_FIPS", "COUNTY_CAT")
msa <- msa %>%
  filter(Met_Mic == "Metropolitan Statistical Area",
         STATE != "Puerto Rico") %>%
  mutate(FIPS = paste0(ST_FIPS, CTY_FIPS)) %>%
  dplyr::select(FIPS, ST_FIPS, CTY_FIPS, COUNTY, STATE, CBSA_FIPS, CBSA_Title)

#Metropolitan statistical area price parities
marpp <- read.csv("./Data/Economic/MARPP/MARPP_MSA_2008_2019.csv")
marpp <- marpp %>%
  mutate(across(where(is.factor), as.character)) %>%
  dplyr::select(GeoFIPS, GeoName, Description, Unit, X2018) %>%
  mutate(GeoFIPS = trimws(GeoFIPS, which="both")) %>%
  filter(!is.na(X2018),
         GeoFIPS != "00999")
colnames(marpp)[5] <- "RPP"
marpp <- marpp %>%
  mutate(RPP = as.numeric(RPP))

#joining marpp to county
marpp <- merge(msa, marpp, by.x="CBSA_FIPS", by.y="GeoFIPS")
marpp <- marpp %>%
  dplyr::select(FIPS, COUNTY, STATE, Description, Unit, RPP)

#Portion area regional price parities
parpp <- read.csv("./Data/Economic/PARPP/PARPP_PORT_2008_2019.csv")
parpp <- parpp %>%
  dplyr::select(GeoFIPS, GeoName, Description, Unit, X2018) %>%
  mutate(GeoFIPS = as.character(GeoFIPS)) %>%
  mutate(GeoFIPS = trimws(GeoFIPS, which="both")) %>%
  mutate(ST_FIPS = substr(GeoFIPS, 1, 2)) %>%
  filter(!is.na(X2018),
         GeoName %in% parpp$GeoName[grep("Nonmetro", parpp$GeoName)])
colnames(parpp)[5] <- "RPP"
parpp <- parpp %>%
  mutate(RPP = as.numeric(RPP))

#Nonmetropolitan Counties
cty <- st_drop_geometry(read_sf("./Data/Census/tl_2019_us_county.shp", quiet=T))
parpp <- merge(cty %>%
                 filter(!GEOID %in% marpp$FIPS,
                        STATEFP %in% unique(msa$ST_FIPS)) %>%
                 dplyr::select(GEOID, NAME, NAMELSAD, STATEFP), parpp, by.x="STATEFP", by.y="ST_FIPS") 
parpp <- parpp %>%
  mutate(across(where(is.factor), as.character)) %>%
  mutate(GeoName = str_replace(GeoName, " \\(Nonmetropolitan Portion\\)", "")) %>%
  dplyr::select(GEOID, NAMELSAD, GeoName, Description, Unit, RPP)
colnames(parpp) <- colnames(marpp)

#All counties join
rpp <- bind_rows(marpp, parpp) %>%
  mutate(Description = trimws(Description, which="both"))
write.csv("./Data/Economic/Regional Price Parities.csv")

#Median household income, 2019
#County Level
st_all <- st_drop_geometry(read_sf("./Data/Census/cb_2019_us_state_20m.shp", quiet=T)) %>%
  filter(!NAME %in% c("Puerto Rico"))

mhi_cty <- list()
for (i in sort(unique(st_all$NAME))){
  mhi_cty[[i]] <- acs5y(2019, state = i, cenunit = "County", table = "B19019") 
}
mhi_cty <- bind_rows(mhi_cty)
mhi_cty <- mhi_cty %>% mutate(FIPS = substr(GEOID, 8, 12))

mhi_cols <- acs5y(2019, table="B19019")
mhi_cty <- mhi_cty %>%
  pivot_longer(cols = starts_with("B19"), names_to="TableID", values_to="MHI")
mhi_cty <- merge(mhi_cty, mhi_cols, by="TableID")
mhi_cty <- mhi_cty %>%
  mutate(TableTitle = gsub("-", " ", TableTitle))

#HH type and size
hh_cty <- list()
for (i in sort(unique(st_all$NAME))[16:51]){
  hh_cty[[i]] <- acs5y(2019, state = i, cenunit = "County", table = "B11016") 
}
hh_cty <- bind_rows(hh_cty)
hh_cty <- hh_cty %>% mutate(FIPS = substr(GEOID, 8, 12))

hh_cols <- acs5y(2019, table="B11016")
hh_cty <- hh_cty %>%
  pivot_longer(cols = starts_with("B11"), names_to="TableID", values_to="HH_Size")
hh_cty <- merge(hh_cty, hh_cols, by="TableID")
hh_cty <- hh_cty %>%
  filter(TableTitle != "Total:") %>%
  mutate(TableTitle = paste0(gsub("-", " ", TableTitle), "s"),
         TableTitle = ifelse(TableTitle %in% c("Family households:s", "Nonfamily households:s"), "Total:", TableTitle),
         HH_Type = ifelse(TableID %in% c("B11016_002", "B11016_003", "B11016_004", "B11016_005", "B11016_006", "B11016_007", "B11016_008"), "Fam_HH", "Nonfam_HH")) %>%
  pivot_wider(id_cols = -c(TableID, HH_Type, HH_Size), names_from=HH_Type, values_from=HH_Size)%>%
  mutate(Fam_HH = ifelse(is.na(Fam_HH), 0, Fam_HH),
         Tot_HH = Fam_HH + Nonfam_HH)

#joining MHI and households
mhi_cty <- merge(hh_cty %>%
                  dplyr::select(FIPS, TableTitle, Tot_HH, Fam_HH, Nonfam_HH),
                mhi_cty %>%
                  dplyr::select(FIPS, TableTitle, MHI),
                by=c("FIPS", "TableTitle"))

#Loading US MHI
us_mhi <- read.csv("./Data/Economic/US_MHI.csv") %>%
  mutate(across(where(is.factor), as.character))
colnames(us_mhi) <- c("TableTitle", "US_MHI", "MOE")
us_mhi <- us_mhi %>%
  mutate(TableTitle = ifelse(TableTitle != "Total:", substr(gsub("-", " ", TableTitle), 5, nchar(TableTitle)), "Total:"),
         US_MHI = as.numeric(gsub(",", "", US_MHI)))

#joining with US MHI
mhi_cty <- merge(mhi_cty, us_mhi %>%
                  dplyr::select(-MOE),
                by="TableTitle")

#joining with regional price parities
mhi_cty <- merge(mhi_cty, 
                rpp %>%
                  filter(Description == "RPPs: Goods") %>%
                  mutate(RPP = RPP/100) %>%
                  dplyr::select(-Unit),
                by="FIPS") %>%
  mutate(MHI_ADJ = MHI/RPP)

mhi_fix <- mhi_cty %>%
  filter(TableTitle != "Total:",
         !is.na(MHI_ADJ)) %>%
  mutate(RelMHI = MHI_ADJ/US_MHI) %>%
  mutate(wRelMHI = RelMHI*Tot_HH) %>%
  group_by(FIPS) %>%
  summarise(Tot_HH = sum(Tot_HH),
            wRelMHI = sum(wRelMHI)) %>%
  ungroup() %>%
  mutate(RelMHIFix = wRelMHI/Tot_HH)

#Median household income, 2019
#Tract Level
mhi <- list()
for (i in sort(unique(st_all$NAME))[36]){
  mhi[[i]] <- acs5y(2019, state = i, cenunit = "Tract", table = "B19019") 
}
mhi <- bind_rows(mhi)
mhi <- mhi %>% mutate(TractID = substr(GEOID, 8, 18))

mhi_cols <- acs5y(2019, table="B19019")
mhi <- mhi %>%
  pivot_longer(cols = starts_with("B19"), names_to="TableID", values_to="MHI")
mhi <- merge(mhi, mhi_cols, by="TableID")
mhi <- mhi %>%
  mutate(TableTitle = gsub("-", " ", TableTitle))

#HH type and size
hh <- list()
for (i in sort(unique(st_all$NAME))){
  hh[[i]] <- acs5y(2019, state = i, cenunit = "Tract", table = "B11016") 
}
hh <- bind_rows(hh)
hh <- hh %>% mutate(TractID = substr(GEOID, 8, 18))
hh_cols <- acs5y(2019, table="B11016")
hh <- hh %>%
  pivot_longer(cols = starts_with("B11"), names_to="TableID", values_to="HH_Size")
hh <- merge(hh, hh_cols, by="TableID")
hh <- hh %>%
  filter(TableTitle != "Total:") %>%
  mutate(TableTitle = paste0(gsub("-", " ", TableTitle), "s"),
         TableTitle = ifelse(TableTitle %in% c("Family households:s", "Nonfamily households:s"), "Total:", TableTitle),
         HH_Type = ifelse(TableID %in% c("B11016_002", "B11016_003", "B11016_004", "B11016_005", "B11016_006", "B11016_007", "B11016_008"), "Fam_HH", "Nonfam_HH")) %>%
  pivot_wider(id_cols = -c(TableID, HH_Type, HH_Size), names_from=HH_Type, values_from=HH_Size)%>%
  mutate(Fam_HH = ifelse(is.na(Fam_HH), 0, Fam_HH),
         Tot_HH = Fam_HH + Nonfam_HH)

#joining MHI and households
mhi_hh <- merge(hh %>%
                  dplyr::select(TractID, TableTitle, Tot_HH, Fam_HH, Nonfam_HH),
                mhi %>%
                  dplyr::select(TractID, TableTitle, MHI),
                by=c("TractID", "TableTitle"))

#joining with US MHI
mhi_hh <- merge(mhi_hh, us_mhi %>%
                  dplyr::select(-MOE),
                by="TableTitle")

#joining with regional price parities
mhi_hh <- merge(mhi_hh %>%
                  mutate(FIPS = substr(TractID, 1, 5)), 
                rpp %>%
                  filter(Description == "RPPs: Goods") %>%
                  mutate(RPP = RPP/100) %>%
                  dplyr::select(-Unit),
                by="FIPS") %>%
  mutate(MHI_ADJ = MHI/RPP)

mhi2 <- mhi_hh %>%
  filter(TableTitle != "Total:") %>%
  mutate(RelMHI = MHI_ADJ/US_MHI) %>%
  mutate(wRelMHI = RelMHI*Tot_HH) %>%
  group_by(TractID) %>%
  summarise(Tot_HH = sum(Tot_HH, na.rm=T),
            wRelMHI = sum(wRelMHI, na.rm=T)) %>%
  ungroup() %>%
  mutate(RelMHI = wRelMHI/Tot_HH)

#joining to tract population counts
pop  <- list()
for (i in sort(unique(st_all$NAME))){
  pop[[i]] <- acs5y(2019, state = i, cenunit = "Tract", table = "B01003") 
}
pop <- bind_rows(pop)
pop <- pop %>% mutate(TractID = substr(GEOID, 8, 18))
pop <- pop %>%
  mutate(TOT_POP = B01003_001) %>%
  dplyr::select(TractID, TOT_POP)

mhi2 <- merge(mhi2, pop %>%
                  dplyr::select(TractID, TOT_POP), by="TractID", all=T)

#Replacing missing mhi with county mhi
mhi2 <- merge(mhi2 %>%
                mutate(FIPS = substr(TractID, 1, 5)),
              mhi_fix %>%
                dplyr::select(FIPS, RelMHIFix), 
              by="FIPS", all=T)
mhi2 <- mhi2 %>%             
  mutate(RelMHI2 = ifelse(TOT_POP != 0 & is.na(RelMHI), RelMHIFix, RelMHI)) %>%
  dplyr::select(-TOT_POP)

write.csv(mhi2, "./Out/Indicators_of_access/US_MHI_RPP_Adjusted.csv")

#subsetting for Appalachian census tracts
app_tract <- st_drop_geometry(read_sf("./Out/SHPs/Appalachia_Tracts.shp"))
mhi2 <- mhi2 %>% 
  dplyr::select(-X) %>%
  mutate(TractID = ifelse(nchar(TractID) == 10, paste0("0", as.character(TractID)), as.character(TractID))) %>%
  filter(TractID %in% app_tract$GEOID)
write.csv(mhi2, "./Out/Indicators_of_access/MHI_RPP_Adjusted.csv")
```

# Vehicle- Percent of Households with No Vehicles

This script calculates household vehicle availability at the census tract level. Two tables are returned, one for Appalachian census tracts alone, one for all US census tracts.

American Community Survey, 2015-2019 estimates

* ACS Table B08201_001: Total households
* ACS Table B08201_002: Households with no vehicles

```{r}
st_all <- st_drop_geometry(read_sf("./Access Data/Appalachia Census SHPs/State TL 2019/cb_2019_us_state_20m.shp", quiet=T)) %>%
  filter(!NAME %in% c("Puerto Rico"))

vehs <- list()
for (i in sort(unique(st_all$NAME))){
  vehs[[i]] <- acs5y(2019, state = i, cenunit = "Tract", table = "B08201") 
}

vehs <- bind_rows(vehs)
vehs <- vehs %>% mutate(GEOID = substr(GEOID, 8, 18)) %>% 
  dplyr::select(GEOID, B08201_001, B08201_002) 
colnames(vehs) <- c("GEOID", "TOTAL_HH", "TOTAL_HUNV")
vehs <- vehs %>%
  mutate(PCT_HUNV = ifelse(TOTAL_HH !=0, TOTAL_HUNV/TOTAL_HH*100, NA))

write_csv(vehs, "./Out/Indicators_of_access/US_Vehicle_Tract.csv")

#subsetting for Appalachian census tracts
app_tract <- st_drop_geometry(read_sf("./Out/SHPs/Appalachia_Tracts.shp"))
vehs <- vehs %>% 
  filter(GEOID %in% app_tract$GEOID)
write.csv(vehs, "./Out/Indicators_of_access/Vehicle_Tract.csv")
```
